A Standard for the Scholarly Citation of Archaeological Data

### 2017-09-12 analysis/paper/paper.Rmd
most of the figures looking good now

@@

Stodden (2010) surveyed registrants at a top machine learning conference (the Neural Information Processing Systems (NIPS) conference) to identify the factors that affect researchers’ decisions to reveal data (Table \@ref(tab:costsbenefits)). Stodden found that communitarian factors influence scientists to share their data, but private incentives determine when scientists choose not to share their code and data. In the top three reasons not to share, we see a lack of incentives to invest time in data sharing (however, there are many other time-consuming tasks in research, and they get done because clear reward structures are in place); a lack of faith in data citation (which we hope this paper will help with); and a lack of an efficient copyright process. On this last point, Table X shows some of the current norms on this, based on Stodden’s (2009a) detailed legal discussion. See also Stodden’s (2009b) ‘Reproducible Research Standard’, summarised in Figure \@(ref:stodden-rrs) for a convenient and robust copyright schema for typical research products.

@@

```{r fig.cap="Summary of results of email requests for data"}
knitr::include_graphics("../figures/email_sharing_plots_hand_edit.png")
```

@@

The _Journal of Archaeological Science_ has had a ‘data disclosure’ policy since at least 2013 (Editors, 2013) which states that “All data relating to the article must be made available in Supplementary files or deposited in external repositories and linked to within the article”. The data availability rate of 53% reported in our pilot study reflects either weak enforcement of this policy or an incomplete understanding among editors and authors of how to interpret and implement this policy. The prominence of compositional and dating data among the shared data types suggests a least effort strategy by authors sharing data that do not require extensive cleaning and tidying (because the data are generated by instrumented rather than entered by hand, such as artefact attribute data). The size of compositional and dating datasets means that they mostly fit easily into tables in the text or supplementary files, and do not require much effort to organise (e.g. very little metadata is required for others to use these data). The small proportion of papers in our sample using or sharing a scripting language suggests that unlike other fields (Eglen et al. 2017), archaeologists do not appear to recognise code for data analysis as a primary research object to share with the research community.

@@

```{r  fig.cap="Summary of relationships between file types, data types, and data locations"}
plot_data_type <- plot_grid(plot_file_type_location, 
                            plot_data_type_location,
                      labels="AUTO", ncol = 1)
plot_data_type
```

@@

```{r eval = TRUE}
library(tidyverse)
# install.packages("rdatacite")

@@

```{r eval = TRUE}
library(tidyverse)
library(stringr)
# glimpse(dc_arch)

@@

```{r eval = TRUE}
# select cols of interest
xx <- 
dc_arch %>% 
  select( datacentre,
          allocator ,
          date,
          creator,
          description,
          language,
          contributor,
          rights,
          subject,
          format,
          publisher)

@@

```{r first-dates, eval = TRUE}
dates <- 
  dc_arch

@@

 peak_year <- 
   dates %>% 
   group_by(yyyy_date) %>% 
   tally(sort = TRUE) %>% 
   slice(1) %>% 
   pull(yyyy_date)

@@

library(viridis)
date_plot <- 
  dates  %>% 
  filter(!is.na(yyyy_date)) %>% 
  ggplot(aes(yyyy_date,
             fill = datacentre
             )) +
  geom_bar() +
  # guides(fill=FALSE) +
  scale_fill_viridis(discrete = TRUE,
                     breaks = top_data_sources[1:5],
                     labels = top_data_sources[1:5]) +
  theme_new() +
  theme(legend.position = c(0.3, 0.75)) +
  scale_x_continuous(limits = c(1950,2017), expand = c(0,0)) +
  xlab("Year data was collected") +
  ylab("Number of archaeology DOIs\nknown to DataCite") +
  geom_vline(xintercept = peak_year, colour = "red") +
  annotate("text", 
           x = peak_year - 3,
           y = ymax,
           label = peak_year)

# plotly::ggplotly(date_plot2) %>%
#   layout(showlegend = FALSE)

# show plot 
date_plot

# change in each major repo over time
dates  %>% 
  filter(!is.na(yyyy_date)) %>% 
  group_by(yyyy_date, datacentre)  %>% 
  tally() %>% 
  filter(n != 0,
         datacentre %in% top_data_sources[1:9]) %>% 
   ggplot(aes(yyyy_date,
              n
             )) +
  geom_line(size = 1)  +
  scale_color_viridis(discrete = TRUE) +
  scale_y_log10() +
  xlim(c(1990, 2017)) +
  theme_new() +
  facet_wrap(~ datacentre)
```

@@

```{r file-types, eval = TRUE}
# File types

@@

# which group has the highest *proportion* of structured data files?
proportion_structured_data_files <- 
dc_arch %>% 
    mutate(strutured_data = ifelse(grepl("access|mdb|dbf|dbase|accdb|dbase|arcgis|shape|shp|mapinfo|mid|mif|surfer|xls|excel|csv|spreadsheet|spreadsheet", tolower(format)), "has_structured_data", "no")) %>% 
  group_by(datacentre, strutured_data) %>% 
  tally(sort = TRUE) %>% 
  spread(strutured_data, n) %>% 
  mutate(perc_structured = has_structured_data / (has_structured_data + no) * 100) %>% 
  arrange(desc(perc_structured))

@@

```{r rights, eval = TRUE}
rights_summary <- 
  rights %>% 
  mutate(the_text =  toupper(gsub(" |_|\\(|\\)", "-", Var1)),
         the_text = gsub("(\\-)\\1+", "\\1", the_text),
         the_text = gsub("-$", "", the_text)) %>% 
  mutate(cc = case_when(
    grepl("SEMANTICS/OPENACCESS", the_text) ~ "EU Open Access",
    grepl("SEMANTICS/RESTRICTEDACCESS", the_text) ~ "EU Restricted Access",
    grepl("SEMANTICS/CLOSEDACCESS", the_text) ~ "EU Closed Access",
    grepl("CC-BY-[^NC-ND]|CC-BY$|ATTRIBUTION-LICENSE|ATTRIBUTION-[^NON]", the_text) ~ "CC-BY",
    grepl("CC-BY-NC-[^ND]|BY-NC-[^ND]|CC-NC$|ATTRIBUTION-NON-COMMERCIAL-", the_text) ~ "CC-BY-NC",
    grepl("CC-BY-NC-SA-[^ND]|CC-NC-SA$|ATTRIBUTION-NON-COMMERCIAL-SHARE", the_text) ~ "CC-BY-NC-SA",
    grepl("CC-BY-NC-ND|BY-NC-ND|ATTRIBUTION-NONCOMMERCIAL-NO-DERIVATIVE|ATTRIBUTION-NONCOMMERCIAL-NODERIVATIVES|ATTRIBUTION-NONCOMMERCIAL-NODERIVS|NAMENSNENNUNG-NICHTKOMMERZIELL-KEINEBEARBEITUNG|NAMENSNENNUNG-KEINE-KOMMERZIELLE-NUTZUNG", the_text) ~ "CC-BY-NC-ND",
    grepl("ADS|ARCHAEOLOGYDATASERVICE", the_text) ~ "ADS",
    grepl("CC0|CC-0|ZERO", the_text) ~ "CC-0"
  )) %>% 
  select(cc, n) %>% 
  filter(!is.na(cc)) %>% 
  group_by(cc) %>% 
  tally(sort = TRUE) %>% 
  mutate(perc = round(nn / sum(nn) * 100, 0))

@@

Many of the records are grey literature such as unpublished reports, produced by commercial consultants or research projects. Of the `r prettyNum(number_of_records_with_file_type, big.mark = ",")` (`r perc_of_records_with_file_type`%) records that include information about the file types in the deposit, `r perc_only_pdf`% (n = `r prettyNum(n_only_pdf, big.mark = ",")`) contain only PDFs. Structured data files such as spreadsheets, databases or shapefiles are found either alone or together with other file types (typically PDFs) in `r perc_structured_data_file`% (n = `r prettyNum(n_structured_data_file, big.mark = ",")`) records. The highest proportion of records with structured data files are found in the DANS-EASY repository (n = `r prettyNum(n_struc_in_dans, big.mark = ",")`, `r perc_struc_in_dans`%), followed by the ADS with `r perc_struc_in_ads`% (n = `r n_struc_in_ads`). Note that we do not have any file format information for all the records from some smaller repositories in this sample, so it is possible that the true proportion of structured files in these records is higher.

@@

```{r datacite-plots, eval = TRUE}
datacentre_plot <- 
  datacentre %>% 
  mutate(Var1 = str_wrap(Var1, 20)) %>% 
  slice(1:10) %>% 
ggplot(aes(reorder(Var1, n),
           n)) +
  geom_col() +
  coord_flip() +
  xlab("Data centre") +
  ylab("") +
  theme_new() +
  theme(axis.text.y  = element_text(size = 8))

@@

# show plots
datacentre_plot
# allocator_plot
# creator_plot
language_plot
# contributor_plot
rights_plot
# publisher_plot

@@

```{r panel-plot, eval = FALSE}
# panel of plots
# combine plots for lang and data center
datacentre_and_langugage_plot <- 
  datacentre_plot + annotation_custom(ggplotGrob(language_plot),
                                      xmin = 1.15, 
                                      xmax = 6.15,
                                      ymin = 10000,
                                      ymax = 40000
  )

# combine plots for lisc and year
rights_and_date_plot <- 
  rights_plot + annotation_custom(ggplotGrob(date_plot),
                                      xmin = 1, 
                                      xmax = 5,
                                      ymin = 10000,
                                      ymax = 40000
  )

library(gridExtra)
lay <- rbind(c(1,2),
             c(1,2))
grid.arrange(datacentre_and_langugage_plot,
             rights_and_date_plot,
             # datacentre_plot,
             # allocator_plot,
             # creator_plot,
             # contributor_plot,
             # publisher_plot,
             # rights_plot,
             # language_plot,
             # date_plot,
             layout_matrix = lay)

@@

```{r eval = TRUE}

datacentre_rights_table[1:11,] %>% 
  ggplot(aes(cc, datacentre)) + 
  geom_point(aes(size = n),
             colour = "grey80") +
  theme_bw() + 
  xlab("") + 
  ylab("") +
  scale_size_continuous(range=c(5,25)) + 
  geom_text(aes(label = n)) +
  theme(legend.position = "none")

# changes over time
library(ggrepel)

datacentre_by_year <- 
dc_arch %>% 
  group_by(yyyy_created, datacentre) %>% 
  tally(sort = TRUE) %>% 
  ungroup() %>% 
  filter(n > 150)

  ggplot(datacentre_by_year, 
         aes(yyyy_created,
             n,
             colour = datacentre)) +
  geom_line(size = 2) +
  theme_bw()  +
  #scale_y_log10() +
  theme(legend.position = "none")  +
  facet_wrap(~ datacentre, scales = "free_y")

@@

In the pilot studies reported above we see three common approaches to sharing data: sending files privately by email, following a personal request; including data as supplementary files submitted for publication with a journal article; and sharing data by depositing files in a trustworthy, public, DOI-issuing repository. The third method, via a public repository, is by far the most widely recommended method both specifically by archaeologists (Kintigh 2006; Kansa et al. 2011; Kansa and Kansa 2013), and in other sciences (cf. Cranston et al. 2014; Penev et al. 2017; Mounce 2014; Thessen & Patterson 2011; Whitlock 2011).	There are many guidelines and conventions from other disciplines that are relevant to sharing archaeological data (ICPSR 2012; Strasser et al. 2012; Wilkinson et al. 2016). Table XXX summarises some of these, and recommendations specifically written for archaeologists can be found in UK Archaeology Data Service (n.d.), ARCHES project. (n.d.) [archaeology data service manuals]. Many of these guidelines may create new technical and practical challenges to archaeologists who are not accustomed to sharing data, and many not want to dedicate resources to sharing data. However, we should not let the perfect be the enemy of the good. In our own work we try to follow these guidelines with no dedicated resources for data sharing, and with an incomplete knowledge of some of the technical details of repository operations. While we recognise that our data sharing pracitses fall short of all the guidelines, we still see value in sharing data imperfectly, rather than not at all.

>>>>

```{r desirable}
tbl <- readr::read_csv("../figures/tbl_rules_of_data_sharing.csv")
knitr::kable(tbl, caption = "Best practices for preparing data for sharing, from Hart et al. (2016) ")
```

@@

Stodden (2010) surveyed registrants at a top machine learning conference (the Neural Information Processing Systems (NIPS) conference) to identify the factors that affect researchers’ decisions to reveal data (Table \@ref(tab:costsbenefits)). Stodden found that communitarian factors influence scientists to share their data, but private incentives determine when scientists choose not to share their code and data. In the top three reasons not to share, we see a lack of incentives to invest time in data sharing (however, there are many other time-consuming tasks in research, and they get done because clear reward structures are in place); a lack of faith in data citation (which we hope this paper will help with); and a lack of an efficient copyright process. On this last point, Table \@ref(tab:desirable) shows some of the current norms on this, based on Stodden’s (2009a) detailed legal discussion. See also Stodden’s (2009b) ‘Reproducible Research Standard’, summarised in Figure \@ref(fig:stodden-rrs) for a convenient and robust copyright schema for typical research products.

@@

```{r fig.cap="Summary of results of email requests for data. A. Responses by year of journal article publication, B. Student status of first author, C. Country location of first author, D. Journal (AAS = Archaeological and Anthropological Sciences, AMANT = American Antiquity, ARCHMTRY = Archaeometry, JAS = Journal of Archaeological Science, JFA = Journal of Field Archaeology, JHE = Journal of Human Evolution, LT = Lithic Technology, SCI = Science."}
knitr::include_graphics("../figures/email_sharing_plots_hand_edit.png")
```

@@

The _Journal of Archaeological Science_ has had a ‘data disclosure’ policy since at least 2013 (Editors, 2013) which states that “All data relating to the article must be made available in Supplementary files or deposited in external repositories and linked to within the article”. The data availability rate of 53% reported in our pilot study reflects either weak enforcement of this policy or an incomplete understanding among editors and authors of how to interpret and implement this policy. The prominence of compositional and dating data among the shared data types suggests a least effort strategy by authors sharing data that do not require extensive cleaning and tidying (because the data are generated by instrumentes rather than entered by hand, such as artefact attribute data). The size of compositional and dating datasets means that they mostly fit easily into tables in the text or supplementary files, and do not require much effort to organise (e.g. very little metadata is required for others to use these data). The small proportion of papers in our sample using or sharing a scripting language suggests that unlike other fields (Eglen et al. 2017), archaeologists do not appear to recognise code for data analysis as a primary research object to share with the research community.

@@

```{r  fig.cap="Summary of relationships between file types, data types, and data locations. A. File types and data locations, B. Data types and data locations. "}
plot_data_type <- plot_grid(plot_file_type_location, 
                            plot_data_type_location,
                      labels="AUTO", ncol = 1)
plot_data_type
```

@@

```{r}
library(tidyverse)
# install.packages("rdatacite")

@@

```{r}
library(tidyverse)
library(stringr)
# glimpse(dc_arch)

@@

```{r}
# select cols of interest
xx <- 
dc_arch %>% 
  select( datacentre,
          allocator ,
          date,
          creator,
          description,
          language,
          contributor,
          rights,
          subject,
          format,
          publisher, 
          yyyy_date, 
          yyyy_created)

@@

```{r first-dates}
dates <- 
  dc_arch

@@

 peak_year <- 
   dates %>% 
   group_by(yyyy_date) %>% 
   tally(sort = TRUE) %>% 
   slice(c(1,4)) %>% 
   pull(yyyy_date)

@@

library(viridis)
date_plot <- 
  dates  %>% 
  filter(!is.na(yyyy_date)) %>% 
  ggplot(aes(yyyy_date,
             fill = datacentre
             )) +
  geom_bar() +
  # guides(fill=FALSE) +
  scale_fill_viridis(discrete = TRUE,
                     breaks = top_data_sources[1:5],
                     labels = top_data_sources[1:5]) +
  theme_new() +
  theme(legend.position = c(0.3, 0.5)) +
  scale_x_continuous(limits = c(1950,2017), expand = c(0,0)) +
  xlab("Year data was collected") +
  ylab("Number of archaeology DOIs\nknown to DataCite") +
  geom_vline(xintercept = peak_year, colour = "red") +
  annotate("text", 
           x = peak_year - 1.5,
           y = ymax,
           label = peak_year,
           size = 3) # +
  # coord_fixed(ratio = 1/1000)

# change in each major repo over time
per_repo <- 
dates  %>% 
  filter(!is.na(yyyy_date)) %>% 
  group_by(yyyy_date, datacentre)  %>% 
  tally() %>% 
  filter(n != 0,
         datacentre %in% top_data_sources[1:9]) %>% 
    mutate(datacentre = str_wrap(datacentre, width = )) %>% 
   ggplot(aes(yyyy_date,
              n
             )) +
  geom_line(size = 1)  +
  scale_color_viridis(discrete = TRUE) +
  scale_y_log10() +
  xlim(c(1990, 2017)) +
  xlab("Trends over time at major archaeological data repositories") +
  theme_new() +
  facet_wrap(~ datacentre)

# show plots
datacite_temporal_plots <- 
 plot_grid(date_plot, 
           per_repo,
           labels="AUTO", 
           ncol = 1,
           align = "hv",
           axis = "lr")

save_plot("../figures/datacite_temporal_plots.png", 
          datacite_temporal_plots,
          base_height = 8)
```

```{r datacite-temporal-summary, fig.cap = "Change over time in archaeology datasets known to DataCite. A. Counts of datasets per year by the major data centres, red lines indicate major increases in the number of datasets in repositories, Note that the date on the horizontal axis indicates when the data was collected, not when it was deposited in the data centre, B. Temporal trends in the number of DOIs at individual data centres."}
knitr::include_graphics("../figures/datacite_temporal_plots.png")
```

@@

```{r file-types}
# File types

@@

# which group has the highest *proportion* of structured data files?
structured_data_files <- 
dc_arch %>% 
    mutate(strutured_data = ifelse(grepl("access|mdb|dbf|dbase|accdb|dbase|arcgis|shape|shp|mapinfo|mid|mif|surfer|xls|excel|csv|spreadsheet|spreadsheet", tolower(format)), "has_structured_data", "no"))

proportion_structured_data_files <- 
  structured_data_files %>% 
  group_by(datacentre, strutured_data) %>% 
  tally(sort = TRUE) %>% 
  spread(strutured_data, n) %>% 
  mutate(perc_structured = has_structured_data / (has_structured_data + no) * 100) %>% 
  arrange(desc(perc_structured))

@@

```{r, fig.cap = "Relationships between data centres and file formats. The ten most frequently found file formats are shown here, in the data centres with these files from the top ten data centres by number of DOIs. Open file formats for structured data, such as CSV, are not common."}
library(tidyr)

structured_data_files_per_file <- 
dc_arch %>% 
    mutate(has_data = ifelse(grepl("access|mdb|dbf|dbase|accdb|dbase|arcgis|shape|shp|mapinfo|mid|mif|surfer|xls|excel|csv|spreadsheet|spreadsheet|pdf", tolower(format)), "has_data", "no")) %>% 
  filter(has_data == "has_data") %>% 
  mutate(format_simple = gsub("application|text|image|\\(|\\)", "", format)) %>% 
  separate(format_simple, 
           sapply(1:50, function(i) paste0('f', i)),
           sep = "/|,|:|;|-|\\.| ") %>% 
  select(datacentre, dataset_id, f1:f50) %>% 
  gather(key = variable, 
         value = value, 
         -datacentre,
         -dataset_id,
         na.rm = TRUE) %>% 
  select(-variable) %>% 
  filter(value != "") %>% 
  mutate(format = str_trim(tolower(value))) %>% 
  filter(grepl("[a-z]+", format)) %>% 
  select(-value) %>% 
  mutate(format = gsub('ms', "", format)) %>% 
  mutate(format = if_else(format %in% c("xls", "xlsx", "excel"),
                          "xls or xlsx", 
                          format)) %>% 
  filter(format != "") %>% 
  distinct(datacentre, dataset_id, format)

# table(structured_data_files_per_file$datacentre)

top_file_formats <- 
  structured_data_files_per_file %>% 
  group_by(format) %>% 
  tally(sort = TRUE) %>% 
  slice(1:20)

structured_data_files_per_file_tally <- 
structured_data_files_per_file %>% 
  group_by(format, datacentre) %>% 
  tally(sort = TRUE) %>% 
  filter(format %in% top_file_formats$format) %>% 
  filter(datacentre %in% top_data_sources)

plot_structured_data_files_per_file_tally <- 
ggplot(structured_data_files_per_file_tally,
       aes(str_wrap(datacentre, 20), 
           format)) +
  geom_point(aes(size = n), 
             shape = 16, 
             #alpha = 0.4,
             colour = viridis(3)[3]) +
  geom_text(aes(label = prettyNum(n, big.mark = ",")), size = 3) +
  scale_size_continuous(range=c(2,25), guide = FALSE) +
    theme_bw() +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, 
                                   hjust = 1)) +
  coord_flip() +
  xlab("Data centre") +
  ylab("File format")

ggsave("../figures/plot_structured_data_files_per_file_tally.png",
       plot_structured_data_files_per_file_tally,
       width = 10)

knitr::include_graphics("plot_structured_data_files_per_file_tally")
```

```{r rights}
rights_summary <- 
  rights %>% 
  mutate(the_text =  toupper(gsub(" |_|\\(|\\)", "-", Var1)),
         the_text = gsub("(\\-)\\1+", "\\1", the_text),
         the_text = gsub("-$", "", the_text)) %>% 
  mutate(cc = case_when(
    grepl("SEMANTICS/OPENACCESS", the_text) ~ "EU Open Access",
    grepl("SEMANTICS/RESTRICTEDACCESS", the_text) ~ "EU Restricted Access",
    grepl("SEMANTICS/CLOSEDACCESS", the_text) ~ "EU Closed Access",
    grepl("CC-BY-[^NC-ND]|CC-BY$|ATTRIBUTION-LICENSE|ATTRIBUTION-[^NON]", the_text) ~ "CC-BY",
    grepl("CC-BY-NC-[^ND]|BY-NC-[^ND]|CC-NC$|ATTRIBUTION-NON-COMMERCIAL-", the_text) ~ "CC-BY-NC",
    grepl("CC-BY-NC-SA-[^ND]|CC-NC-SA$|ATTRIBUTION-NON-COMMERCIAL-SHARE", the_text) ~ "CC-BY-NC-SA",
    grepl("CC-BY-NC-ND|BY-NC-ND|ATTRIBUTION-NONCOMMERCIAL-NO-DERIVATIVE|ATTRIBUTION-NONCOMMERCIAL-NODERIVATIVES|ATTRIBUTION-NONCOMMERCIAL-NODERIVS|NAMENSNENNUNG-NICHTKOMMERZIELL-KEINEBEARBEITUNG|NAMENSNENNUNG-KEINE-KOMMERZIELLE-NUTZUNG", the_text) ~ "CC-BY-NC-ND",
    grepl("ADS|ARCHAEOLOGYDATASERVICE", the_text) ~ "ADS",
    grepl("CC0|CC-0|ZERO", the_text) ~ "CC-0"
  )) %>% 
  select(cc, n) %>% 
  filter(!is.na(cc)) %>% 
  group_by(cc) %>% 
  tally(sort = TRUE) %>% 
  mutate(perc = round(nn / sum(nn) * 100, 0))

@@

Many of the records are grey literature such as unpublished reports, produced by commercial consultants or research projects. Of the `r prettyNum(number_of_records_with_file_type, big.mark = ",")` (`r perc_of_records_with_file_type`%) records that include information about the file types in the deposit, `r perc_only_pdf`% (n = `r prettyNum(n_only_pdf, big.mark = ",")`) contain only PDFs. Structured data files such as spreadsheets, databases or shapefiles are found either alone or together with other file types (typically PDFs) in `r perc_structured_data_file`% (n = `r prettyNum(n_structured_data_file, big.mark = ",")`) records. The highest proportion of records with structured data files are found in the DANS-EASY repository (n = `r prettyNum(n_struc_in_dans, big.mark = ",")`, `r perc_struc_in_dans`%), followed by the ADS with `r perc_struc_in_ads`% (n = `r n_struc_in_ads`). Note that file format information is not present for all records in this sample, and some of the smaller repositories do not supply this information to DataCite at all, so it is possible that the true proportion of structured files in these records is higher than obsevered here.

@@

```{r datacite-plots}
datacentre_plot <- 
  datacentre %>% 
  mutate(Var1 = str_wrap(Var1, 20)) %>% 
  slice(1:10) %>% 
ggplot(aes(reorder(Var1, n),
           n)) +
  geom_col() +
  coord_flip() +
  xlab("Data centre") +
  ylab("") +
  theme_new() +
  theme(axis.text.y  = element_text(size = 8))

@@

# show plots
# datacentre_plot
# allocator_plot
# creator_plot
# language_plot
# contributor_plot
# rights_plot
# publisher_plot

# Wow, we can nest cowplot::plot_grid in gridExtra::grid.arrange !! amazing and easy!
datacite_plots <- 
arrangeGrob(datacentre_plot, 
             plot_grid(language_plot, 
                       rights_plot, 
                       ncol=1, 
                       align="v",
                       axis = "lr"), 
             layout_matrix = rbind(c(1,2),
                                   c(1,2)),
             ncol = 2)

save_plot("../figures/datacite_plots.png", 
          grid.draw(datacite_plots),
          base_height = 8)
```

```{r datacite-summary, fig.cap = "Distributions of data centers, languages (en = English, nl = Dutch, swe = Swedish, de = German, it = Italian) and licenses (see text for details) among archaeology datasets known to DataCite. Horizontal axis is the number of DOIs."}
knitr::include_graphics("../figures/datacite_plots.png")
```

@@

```{r}

plot_datacentre_rights_table <- 
datacentre_rights_table[1:11,] %>% 
  ggplot(aes(cc, str_wrap(str_trunc(datacentre, side = "right", 35), 15))) + 
  geom_point(aes(size = n),
             colour = "grey80",
             shape = 16) +
  theme_bw() + 
  xlab("") + 
  ylab("") +
  scale_size_continuous(range=c(7,25)) + 
  geom_text(aes(label = prettyNum(n, big.mark = ","))) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

ggsave("../figures/datacentre_rights_table.png", 
       plot_datacentre_rights_table,
       height = 7)

# changes over time
# library(ggrepel)
# 
# datacentre_by_year <- 
# dc_arch %>% 
#   group_by(yyyy_created, datacentre) %>% 
#   tally(sort = TRUE) %>% 
#   ungroup() %>% 
#   filter(n > 150)
# 
# 
#   ggplot(datacentre_by_year, 
#          aes(yyyy_created,
#              n,
#              colour = datacentre)) +
#   geom_line(size = 2) +
#   theme_bw()  +
#   scale_y_log10() +
#   theme(legend.position = "none")  +
#   facet_wrap(~ datacentre, scales = "free_y")

@@

```{r fig.cap = "Summary of licenses used by data centers with archaeological data"}
knitr::include_graphics("../figures/datacentre_rights_table.png")
```

@@

In the pilot studies reported above we see three common approaches to sharing data: sending files privately by email, following a personal request; including data as supplementary files submitted for publication with a journal article; and sharing data by depositing files in a trustworthy, public, DOI-issuing repository. The third method, via a public repository, is by far the most widely recommended method both specifically by archaeologists (Kintigh 2006; Kansa et al. 2011; Kansa and Kansa 2013), and in other sciences (cf. Cranston et al. 2014; Penev et al. 2017; Mounce 2014; Thessen & Patterson 2011; Whitlock 2011).	There are many guidelines and conventions from other disciplines that are relevant to sharing archaeological data (ICPSR 2012; Strasser et al. 2012; Wilkinson et al. 2016). Table \@ref(tab:desirable) summarises some of these, and recommendations specifically written for archaeologists can be found in UK Archaeology Data Service (n.d.), ARCHES project. (n.d.) [archaeology data service manuals]. Many of these guidelines may create new technical and practical challenges to archaeologists who are not accustomed to sharing data, and many not want to dedicate resources to sharing data. However, we should not let the perfect be the enemy of the good. In our own work we try to follow these guidelines with no dedicated resources for data sharing, and with an incomplete knowledge of some of the technical details of repository operations. While we recognise that our data sharing pracitses fall short of all the guidelines, we still see value in sharing data imperfectly, rather than not at all.
